<span style="color: red"> THELANCET-D-15-02346 </span>

Open access raw data, tidy data, code book and statistical inference for the journal article:

Living longer with disability in China and G20 nations: a quantitative model for the Sustainable Development Goal for health post-2015. A reproducible research
==========================================================================================

## Accepted at the Lancet-Chinese Academy of Medical Sciences (CAMS) Health Summit, Beijing, China, Oct 30–31, 2015


### Dr Stefano Olgiati, PhD, MS (1), Prof. Alessandro Danovi, MS (2)

1.  Author Responsible For Correspondence: University of Bergamo <www.unibg.it>, Department of Quantitative Methods, via dei Caniana 2, Bergamo (ITA) 24129; tel  +39 035 20 52 638; fax +39 035 20 52 549; email <stefano.olgiati@unibg.it>
2. University of Bergamo, Faculty of Economics, Bergamo (ITA) 24129

<span style="color: red">
__NOTICE TO FELLOW RESEARCHERS:___ 
___1) THE WHO AS OF JULY 18, 2015, HAS UPDATED SOME DATABASES UTILIZED AS RAW DATA FOR THE TIDY DATASETS AND THE STATISTICAL INFERENCE. ACCORDINGLY, AFTER RUNNING THE PRESENT PROGRAM SOME OF THE NUMERIC FINDINGS MAY SLIGHTLY DIFFER  FROM THE FINDINGS REPORTED IN THE JOURNAL ARTICLE THELANCET-D-15-02346 (DOI not yet released). THE AUTHORS HAVE CHOSEN NOT TO "FREEZE" THE ORIGINAL DATA SOURCE BUT TO ALLOW AUTOMATIC UPDATING OF THE RESEARCH FINDINGS IN THE HTML DOCUMENT. THIS CAN BE VERIFIED BY THE > dateDownloaded CODE IN R AND TRACED BY SUBSEQUENT UPDATES IN THE OA GITHUB REPOSITORY. THIS IS A FURTHER ADVANTAGE IN TERMS OF TRANSPARENCY OF RESEARCH REPRODUCIBILITY;___

__2)AS OF OCTOBER 2, 2015 THE UNIVERSITY OF WASHINGTON GBDx HAS PERMANENTLY MOVED THE URL. THE DOWNLOAD CODE HAS BEEN UPDATED AND THE MESSAGE FROM THE OLD URL UPLOADED IN GITHUB FOR TRANSPARENCY AND EVIDENCE__

__3)AS OF OCTOBER 2, 2015 SOFTWARE UPDATE TO MAC OS X EL CAPITAN 10.11 HAS REQUIRED TO ADD THE METHOD "curl" TO download.file function in R__

__THANK YOU. THE AUTHORS__
</span>

```{r echo=FALSE, results='hide'}
library(data.table)
library(car) # For computation of variance inflation factors VIF
library(scatterplot3d) 
```

### Raw Data
We utilized open access data from:

1. the Institute for Health Metrics and Evaluation (IHME) Global Burden of Disease Study 2010 [GHDx]("http://ghdx.healthmetricsandevaluation.org/sites/default/files/record-attached-files/IHME_GBD_2010_HALE_BY_COUNTRY_1990_2010_Y2010M02D23.CSV");

2. the World Health Organization’s [Data Repository on Health Expenditure]("http://apps.who.int/gho/athena/data/xmart.csv?target=GHO/WHS7_156,WHS7_105,WHS7_104,WHS7_108&profile=xmart&filter=COUNTRY:*;REGION:AFR;REGION:AMR;REGION:SEAR;REGION:EUR;REGION:EMR;REGION:WPR").

### Code Book
The following variables and definitions, based on “The Global Burden of Disease: Generating Evidence, Guiding Policy” and the World Health Organization’s Data Repository on Health Expenditure, were adopted when extracting the tidy data set from the raw data:

* Life Expectancy at birth: Life Expectancy at birth is the average number of years that a person can be expected to live from age 0-1;

* Health Adjusted Life Expectancy (HALE) at birth: Healthy Life Expectancy at birth is the average number of years that a person can be expected to live in “full health” from age 0-1, by taking into account years lived in less than full health with disease and/or non-fatal injury ;

* Years Lived with Disability (YLD): The Years Lived with Disability is the difference between LE and HALE, where LE is always greater than or equal to HALE. It is equivalent to the number of years that people live with a health condition or its consequences or in less than full health with disease and/or non-fatal injury;

* Total Health Expenditure (THE): WHS7_105 Per capita total expenditure on health (PPP int.$) includes public (government, external resources for health from international programs and social security) and private expenditure (private pre-paid plans and out-of-pocket). For the purpose of international comparisons, THE is expressed in International Purchase Power Parity United States Dollars (USD PPP) which, according to the World Bank, is the number of units of a country’s currency required to buy the same amounts of goods and services in the domestic market as one US Dollar would buy in the United States.

### Raw Data
Downloading data from the repositories

```{r, cache=TRUE}

# Getting raw data from the IHME Global Burden of Disease Study 2010
fileURL.GBD <- "http://ghdx.healthdata.org/sites/default/files/record-attached-files/IHME_GBD_2010_HALE_BY_COUNTRY_1990_2010_Y2010M02D23.CSV"

# Getting raw data from the WHO Health Expenditure Repository
fileURL.WHO <- "http://apps.who.int/gho/athena/data/xmart.csv?target=GHO/WHS7_156,WHS7_105,WHS7_104,WHS7_108&profile=xmart&filter=COUNTRY:*;REGION:AFR;REGION:AMR;REGION:SEAR;REGION:EUR;REGION:EMR;REGION:WPR"

download.file(fileURL.GBD, destfile="GBD.csv", method="curl")
download.file(fileURL.WHO, destfile="WHO.csv", method="curl")

dateDownloaded <- date()
dateDownloaded
```

Reading raw data in .csv format
````{r}
# Raw Data
raw.data.GBD <- read.csv("GBD.csv")
raw.data.WHO <- read.csv("WHO.csv")
names(raw.data.GBD)
names(raw.data.WHO)
dim(raw.data.GBD)
dim(raw.data.WHO)

# Data Tables
attach(raw.data.GBD)
attach(raw.data.WHO)

DT.GBD <- data.table(raw.data.GBD)
DT.WHO <- data.table(raw.data.WHO)

tables()
```


### Tidy Data Set

Manipulating data to generate tidy datasets
```{r}
# ISO3 codes of G20 Countries for 1990/95 - 2010
# Argentina, Australia, Brazil, Canada, China, India, Indonesia, Japan, Mexico, Russia, Saudi Arabia, South Africa, South Korea, Turkey, the United Kingdom and the United States, European Union (EU)

ISO3 <- c("ARG", "AUS", "BRA", "CAN","CHN","IND","IDN","JPN","MEX","RUS","SAU","ZAF","KOR","TUR","USA", "AUT","BEL","BGR","HRV","CYP","CZE","DNK","EST","FIN","FRA","DEU","GRC","HUN","IRL","ITA","LVA","LTU","LUX","MLT","NLD","POL","PRT","ROU","SVK","SVN","ESP","SWE","GBR")

length(ISO3)

data.GBD.1990 <- DT.GBD[DT.GBD$year == "1990" & DT.GBD$age_name == "0-1 years" & DT.GBD$sex_name == "Both" & DT.GBD$iso3 %in% ISO3 , ]
data.GBD.1990.ordered <- data.GBD.1990[order(iso3), ]
data.GBD.1990.ordered

data.GBD.2010 <- DT.GBD[DT.GBD$year == "2010" & DT.GBD$age_name == "0-1 years" & DT.GBD$sex_name == "Both" & DT.GBD$iso3 %in% ISO3 , ]
data.GBD.2010.ordered <- data.GBD.2010[order(iso3), ]
data.GBD.2010.ordered

data.WHO.1995 <- DT.WHO[DT.WHO$YEAR == "1995" & DT.WHO$GHO == "WHS7_105" & DT.WHO$Numeric != "NA" & DT.WHO$COUNTRY %in% ISO3,  ]
data.WHO.1995.ordered <- data.WHO.1995[order(COUNTRY), ]
data.WHO.1995.ordered

data.WHO.2010 <- DT.WHO[DT.WHO$YEAR == "2010" & DT.WHO$GHO == "WHS7_105" & DT.WHO$Numeric != "NA" & DT.WHO$COUNTRY %in% ISO3,  ]
data.WHO.2010.ordered <- data.WHO.2010[order(COUNTRY), ]
data.WHO.1995.ordered

data.1990 <- data.frame(country=data.GBD.1990.ordered$iso3, le=data.GBD.1990.ordered$le, hale=data.GBD.1990.ordered$hale, the=data.WHO.1995.ordered$Numeric)

data.2010 <- data.frame(country=data.GBD.2010.ordered$iso3, le=data.GBD.2010.ordered$le, le_ui = data.GBD.2010.ordered$le_ui, hale=data.GBD.2010.ordered$hale, hale_ui = data.GBD.2010.ordered$hale_ui, the=data.WHO.2010.ordered$Numeric)


# Years Lived with Disease in G20 Countries for 1990 - 2010
yld.1990 <- data.1990$le - data.1990$hale
yld.2010 <- data.2010$le - data.2010$hale

# Tidy Dataset
tidy.data <- data.frame(country=data.2010$country, the_1995=data.1990$the, the_2010=data.2010$the, le_1990=data.1990$le,  le_2010=data.2010$le, le_ui_2010=data.2010$le_ui, hale_ui_2010=data.2010$hale_ui, hale_1990 = data.1990$hale, hale_2010 = data.2010$hale, yld_1990=yld.1990, yld_2010=yld.2010)

attach(tidy.data)

# Saving Tidy Dataset to .csv file
write.csv(tidy.data, "YLDOutput.csv")
```

### Background
```{r}
ID <- "CHN"

# Data for background and rationale
tidy.data$le_2010[tidy.data$country == ID]
tidy.data$le_1990[tidy.data$country == ID]
median(tidy.data$le_2010)
mad(tidy.data$le_2010)

tidy.data$yld_2010[tidy.data$country == ID]
tidy.data$yld_1990[tidy.data$country == ID]
median(tidy.data$yld_2010)
mad(tidy.data$yld_2010)
mean(tidy.data$yld_2010)
var(tidy.data$yld_2010)

tidy.data$hale_2010[tidy.data$country == ID]
tidy.data$hale_1990[tidy.data$country == ID]
median(tidy.data$hale_2010)
mad(tidy.data$hale_2010)
```


### Exploratory Data Analysis
```{r}

# Strip Charts
par(mfrow=(c(1,2)))

strip.data.1 <- data.frame("YEAR 1990"=le_1990, "YEAR 2010"= le_2010)
stripchart(strip.data.1, 
           vertical=TRUE,
           col=c("black", "skyblue"),
           cex=1.5,
           lty = "solid", 
           lwd = 2,
           method="jitter", 
           main="Life Expectancy",
           ylab="Number of Years")
abline(h=c(median(le_1990), median(le_2010)),col=c("black", "skyblue"),
           lty = "solid", 
           lwd = 2) 
points(c(tidy.data$le_1990[tidy.data$country == ID], tidy.data$le_2010[tidy.data$country == ID]), pch=22, col="red", cex = 2, lty = "solid", lwd = 2)
text(c(tidy.data$le_1990[tidy.data$country == ID], tidy.data$le_2010[tidy.data$country == ID]), labels=ID, cex= 1.5, pos=c(4,2))
summary(strip.data.1)

strip.data.4 <- data.frame("YEAR 1990"=the_1995, "YEAR 2010"=the_2010)
stripchart(strip.data.4, 
           vertical=TRUE,
           cex=1.5,
           lty = "solid", 
           lwd = 2,
           col=c("black", "skyblue"),
           method="jitter", 
           main="Per-capita Total\n Health Expenditure",
           ylab="PPP int.$")
abline(h=c(median(the_1995), median(the_2010)),col=c("black", "skyblue"),
           lty = "solid", 
           lwd = 2)
points(c(tidy.data$the_1995[tidy.data$country == ID], tidy.data$the_2010[tidy.data$country == ID]), pch=22, col="red", cex = 2, lty = "solid", lwd = 2)
text(c(tidy.data$the_1995[tidy.data$country == ID], tidy.data$the_2010[tidy.data$country == ID]), labels=ID, cex= 1.5, pos=c(4,2))
summary(strip.data.4)

par(mfrow=(c(1,2)))

strip.data.2 <- data.frame("YEAR 1990"=hale_1990, "YEAR 2010"=hale_2010)
stripchart(strip.data.2, 
           vertical=TRUE,
           col=c("black", "skyblue"),
           cex=1.5,
           lty = "solid", 
           lwd = 2,
           method="jitter", 
           main="Health-Adjusted \n Life Expectancy",
           ylab="Number of Years")
abline(h=c(median(hale_1990), median(hale_2010)),col=c("black", "skyblue"),
           lty = "solid", 
           lwd = 2)
points(c(tidy.data$hale_1990[tidy.data$country == ID], tidy.data$hale_2010[tidy.data$country == ID]), pch=22, col="red", cex = 2, lty = "solid", lwd = 2)
text(c(tidy.data$hale_1990[tidy.data$country == ID], tidy.data$hale_2010[tidy.data$country == ID]), labels=ID, cex= 1.5, pos=c(4,2))
summary(strip.data.2)

strip.data.3 <- data.frame("YEAR 1990"=yld_1990, "YEAR 2010"=yld_2010)
stripchart(strip.data.3, 
           vertical=TRUE,
           col=c("black", "skyblue"),
           cex=1.5,
           lty = "solid", 
           lwd = 2,
           method="jitter", 
           main="Years Lived \n with Disability",
           ylab="Number of Years")
abline(h=c(median(yld_1990), median(yld_2010)),col=c("black", "skyblue"),
           lty = "solid", 
           lwd = 2)
points(c(tidy.data$yld_1990[tidy.data$country == ID], tidy.data$yld_2010[tidy.data$country == ID]), pch=22, col="red", cex = 2, lty = "solid", lwd = 2)
text(c(tidy.data$yld_1990[tidy.data$country == ID], tidy.data$yld_2010[tidy.data$country == ID]), labels=ID, cex= 1.5, pos=c(4,2))
summary(strip.data.3)


# Histograms
par(mfrow=(c(1,3)))
hist(hale_2010, breaks=10, xlab="YEARS", main="HALE")
hist(yld_2010, breaks=10, xlab="YEARS", main="YLD")
hist(the_2010, breaks=10, xlab="PPP int.$", main="THE")

# Normality and skewness
par(mfrow=(c(1,3)))
qqnorm(hale_2010, main="hale_2010")
qqline(hale_2010, col="red")
qqnorm(yld_2010, main="yld_2010")
qqline(yld_2010, col="red")
qqnorm(the_2010, main="the_2010")
qqline(the_2010, col="red")

```

### Model, Multiple Linear Regression and Robustness
````{r}
# Linear regression
model <- lm(log(the_2010)~yld_2010 + hale_2010)

summary(model)

coef(model)
exp(coef(model)["yld_2010"])
exp(coef(model)["hale_2010"])

confint(model)
exp(confint(model))

# VIF
vif(model)

# General Linear Hypothesis Test
library("multcomp")
interact <- glht(model)
interact
summary(interact)

# Analysis of Variance Table
anova(model)

# Prediction
predconf <- predict(model, data.frame(yld_2010 = median(tidy.data$yld_2010), hale_2010 = median(tidy.data$hale_2010)), interval="confidence")
predconf
lower <- exp(predconf[,"lwr"])
fit <- exp(predconf[,"fit"])
upper <- exp(predconf[,"upr"])
mat <- cbind(fit, lower, upper)
df <- data.frame(mat)
df

# Boxplots
par(mfrow=(c(1,3)))

boxplot(the_2010,
        ylab="Per-Capita Total Health Expenditure in PPP int.$")
points(tidy.data$the_2010[tidy.data$country == ID], pch=22, col="red", cex = 2, lty = "solid", lwd = 2)
text(tidy.data$the_2010[tidy.data$country == ID], labels=ID, cex= 1.5, pos=3)

boxplot(hale_2010,
        xlab="YEAR 2010",
        ylab="Health-Adjusted Life Expectancy")
points(tidy.data$hale_2010[tidy.data$country == ID], pch=22, col="red", cex = 2, lty = "solid", lwd = 2)
text(tidy.data$hale_2010[tidy.data$country == ID], labels=ID, cex= 1.5, pos=3)

boxplot(yld_2010,
        ylab="Years Lived with Disability")
points(tidy.data$yld_2010[tidy.data$country == ID], pch=22, col="red", cex = 2, lty = "solid", lwd = 2)
text(tidy.data$yld_2010[tidy.data$country == ID], labels=ID, cex= 1.5, pos=3)


# Summarise the relationship between THE and YLD
par(mfrow=(c(1,1)))

par(mfrow=(c(1,1)))
plot(le_2010, yld_2010,
     cex=1.5,
     lwd = 3,
     xlab="Life Expectancy at Birth",
     ylab="Years Lived with Disability",
     col="skyblue")
points(tidy.data$le_2010[tidy.data$country == ID], tidy.data$yld_2010[tidy.data$country == ID], pch=22, col="red", cex = 2, lty = "solid", lwd = 2)
text(tidy.data$le_2010[tidy.data$country == ID], tidy.data$yld_2010[tidy.data$country == ID], labels=ID, cex= 1, pos=3)

plot(yld_2010, the_2010,
     xlim = c(7, 13),
     ylim = c(0, 10500),
     cex=1.5,
     lwd = 3,
     xlab="Years Lived with Disability",
     ylab="Per-Capita Total Health Expenditure in PPP int.$",
     col="skyblue")
points(tidy.data$yld_2010[tidy.data$country == ID], tidy.data$the_2010[tidy.data$country == ID], pch=22, col="red", cex = 2, lty = "solid", lwd = 2)
text(tidy.data$yld_2010[tidy.data$country == ID], tidy.data$the_2010[tidy.data$country == ID], labels=ID, cex= 1, pos=3)

```

 
